---
title: "HW1"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{HW1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Question 0A: Please let me know your preferred name and preferred pronouns.

John Chen, he/him


Question 0B: Please read the syllabus in its entirety and write, “I have read and understood the entire syllabus,” for this question. If you have any questions about the syllabus (for example, on the homework submission policy or collaboration policy), you may ask them here.

I have read and understood the entire syllabus

Question 0C: If any, please briefly describe the courses you have taken at UW that involve coding (in any language) so far and how familiar/comfortable you are in coding in R. (Note: I do not expect you to have taken any such courses already.)

familar with R markdown, Biostat 515


Question 0D: What do you hope to learn from this course? (It is alright if you leave this blank, but it would be helpful to hear your thoughts.)

Familiar with using and debuging with R and useful techniques for future research projects! Thanks!

```{r setup}
# library(UWBiost561)

plot(1:10)

```

```{r}
source("https://raw.githubusercontent.com/linnykos/561_s2025_public/main/HW1_files/generate_data.R")
set.seed(1)
mean(generate_data(n=10))
#> [1] 5.199818

sample_sizes <- c(1, 2, 3, 5, 10, 100)

par(mfrow = c(2, 3), mar = c(4, 4, 2, 1))  # layout: 2 rows, 3 columns

for (n in sample_sizes) {
  means <- replicate(10000, mean(generate_data(n)))
  hist(
    means,
    breaks = 100,
    main = paste("Sample size:", n),
    xlab = "Value",
    ylab = "Frequency",
    col = "gray",
    border = "black"
  )
}

```
### 2c
The plot shows that when sample size gets bigger, the shape of the sample means becomes more like a normal distribution. This matches what the Central Limit Theorem says. It shows that even if the original data isn’t normal, the average of many samples will be.


### 3a
```{r}
df <- read.csv("https://raw.githubusercontent.com/linnykos/561_s2025_public/main/HW1_files/sea-ad.csv")

head(df)

summary(df)

```

Head functions gives the first 6 rows of all columns in the df, Summary function gives the basic statistics describing this df. 


###3b
```{r}
dim(df)
```

there are 84 rows and 7 columns

### 3c
```{r}
df$Age.at.Death <- gsub("90\\+", "90", df$Age.at.Death)
df$Age.at.Death <- as.numeric(df$Age.at.Death)
hist(df$Age.at.Death)
```
###3d
```{r}
df$Sex <- as.factor(df$Sex)
df$APOE4.Status <- as.factor(df$APOE4.Status)
df$Cognitive.Status <- as.factor(df$Cognitive.Status)
df$Braak <- as.factor(df$Braak)

```

###3e
```{r}
summary(df)
```
it shows na values, general distribution stats for users to look on, instead of just type and length. 

### 3f
```{r}
table(df$Braak, df$Cognitive.Status)
```

### 3g
```{r}
score_bins <- cut(df$Last.CASI.Score,
                  breaks = quantile(
                    df$Last.CASI.Score,
                    probs = seq(0, 1, 0.25),
                    na.rm = TRUE
                  ))

table(score_bins, df$Cognitive.Status)

```
People with lower CASI scores are more likely to have dementia, while those with higher scores are more often categorized as having no dementia. This shows that CASI score is associated with cognitive status.
